---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `testPkgApp`

<!-- badges: start -->
<!-- badges: end -->

The goal of `testPkgApp` is to demonstrate how to build and test a shiny application using `usethis`, `shiny::testServer()`, and `shinytest2`.

## Installation

You don't want to install this package, but you might want to download it as an example (or read through [this post](https://mjfrigaard.github.io/posts/testing-shiny/) to learn about it's contents).

## Run the app

```{r}
#| eval: false
testPkgApp::runShinyApp()
```

# Unit tests 

Check the unit tests for `gg_base()` and `gg_points()` in 

```{r}
#| echo: false
fs::dir_tree("tests/testthat/", regexp = "^tests/testthat/test-gg_")
```


# Shiny server tests 

Check the shiny `testServer()` tests for `mod_var_input_server()` and `mod_display_plot_server()` in

```{r}
#| echo: false
fs::dir_tree("tests/testthat/", regexp = "server.R$")
```




