---
title: "utap"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{utap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "\t",
  eval = TRUE
)
log_maker <- function(size, missing = FALSE) {
  if (size <= 2 & isTRUE(missing)) {
    as.vector(c(TRUE, NA), mode = "logical")
  } else if (size <= 2 & isFALSE(missing)) {
    as.vector(c(TRUE, FALSE), mode = "logical")
  } else if (size > 2 & isTRUE(missing)) {
    rep(c(TRUE, FALSE, NA), length.out = size)
  } else if (size > 2 & isFALSE(missing)) {
    rep(c(TRUE, FALSE), length.out = size)
  }
}
int_maker <- function(size, missing = FALSE) {
  if (size < 3 & isTRUE(missing)) {
    int_raw <- c(as.integer(exp(size)), NA_real_)
    int_vec <- as.vector(int_raw, mode = "integer")
    return(int_vec)
  } else if (size < 3 & isFALSE(missing)) {
    int_raw <- seq.int(from = 1, to = as.integer(exp(size)), length.out = size)
    int_vec <- as.vector(int_raw, mode = "integer")
    return(int_vec)
  } else if (size >= 3 & isTRUE(missing)) {
    adj_size <- round(size * 0.66, digits = 0)
    int_seq <- seq.int(from = 1, to = as.integer(exp(size)), length.out = adj_size)
    int_raw <- rep(c(int_seq, NA_integer_), length.out = size)
    int_vec <- as.vector(int_raw, mode = "integer")
    return(int_vec)
  } else {
    int_raw <- seq.int(from = 1, to = as.integer(exp(size)), length.out = size)
    int_vec <- as.vector(int_raw, mode = "integer")
    return(int_vec)
  }
}
dbl_maker <- function(size, missing = FALSE) {
  if (size < 3 & isTRUE(missing)) {
    dbl_raw <- c(as.double(log(size)), NA_real_)
    dbl_vec <- round(as.vector(dbl_raw, mode = "double"), digits = 3)
    return(dbl_vec)
  } else if (size < 3 & isFALSE(missing)) {
    dbl_raw <- seq.int(from = 0.1, to = as.double(size / 2), length.out = size)
    dbl_vec <- round(as.vector(dbl_raw, mode = "double"), digits = 3)
    return(dbl_vec)
  } else if (size >= 3 & isTRUE(missing)) {
    adj_size <- round(size * 0.33, digits = 0)
    dbl_seq <- seq.int(from = 0.1, to = as.double(size / 2), length.out = adj_size)
    dbl_raw <- rep(c(dbl_seq, NA_real_), length.out = size)
    dbl_vec <- round(as.vector(dbl_raw, mode = "double"), digits = 3)
    return(dbl_vec)
  } else {
    dbl_raw <- seq.int(from = 0.1, to = as.double(size / 2), length.out = size)
    dbl_vec <- round(as.vector(dbl_raw, mode = "double"), digits = 3)
    return(dbl_vec)
  }
}
chr_maker <- function(size, lvls = size/2, missing = FALSE) {
  if (size < lvls) {
    lvls <- size - 1
  }
  if (isTRUE(missing) & size < 3) {
    chr_vec <- as.vector(c("item: 1", NA_character_),
      mode = "character"
    )
  } else if (isFALSE(missing) & size < 3) {
    chr_vec <- as.vector(c("item: 1", "item: 2"),
      mode = "character"
    )
  } else if (isTRUE(missing) & size >= 3) {
    adj_size <- size - 1
    levs <- paste0("item:", as.integer(1:lvls))
    adj_chr <- rep(c(levs, NA_character_), length.out = adj_size)
    nas <- rep(adj_chr, length.out = size)
    chr_vec <- as.vector(c(nas), mode = "character")
  } else {
    levs <- paste0("item:", as.integer(1:lvls))
    chr_raw <- rep(levs, length.out = size)
    chr_vec <- as.vector(c(chr_raw), mode = "character")
  }
  return(chr_vec)
}
fct_maker <- function(size, lvls = size/2, missing = FALSE) {
  if (isTRUE(missing)) {
    levs <- paste0("group ", as.integer(1:lvls))
    nas <- rep(c(levs, NA_character_), length.out = size)
    chr_raw <- as.vector(c(nas), mode = "character")
    fct_vec <- factor(chr_raw,
      levels = unique(sort(chr_raw)),
      ordered = FALSE,
    )
  } else {
    levs <- paste0("group ", as.integer(1:lvls))
    chr_raw <- rep(levs, length.out = size)
    ord_levels <- sort(unique(chr_raw))
    fct_vec <- factor(chr_raw, levels = ord_levels, ordered = FALSE)
  }
  return(fct_vec)
}
ord_maker <- function(size, lvls = size/2, missing = FALSE) {
  if (isTRUE(missing)) {
    levs <- paste0("level ", as.integer(1:lvls))
    nas <- rep(c(levs, NA_character_), length.out = size)
    chr_raw <- as.vector(c(nas), mode = "character")
    ord_vec <- factor(chr_raw,
      levels = unique(sort(chr_raw)),
      ordered = TRUE
    )
  } else {
    levs <- paste0("level ", as.integer(1:lvls))
    chr_raw <- rep(levs, length.out = size)
    ord_levels <- sort(unique(chr_raw))
    ord_vec <- factor(chr_raw, levels = ord_levels, ordered = TRUE)
  }
  return(ord_vec)
}
bin_maker <- function(bin_type, size, missing = FALSE) {
  if (size < 3) {
    size <- 3
    cli::cli_alert_info("size must be >= 3")
  }
  if (isTRUE(missing)) {
    switch(bin_type,
      log = rep(x = c(TRUE, FALSE, NA), length.out = size),
      int = rep(x = c(0L, 1L, NA_integer_), length.out = size),
      chr = rep(
        x = c("item:A", "item:B", NA_character_),
        length.out = size
      ),
      fct = factor(
        rep(
          x = c("group A", "group B", NA_character_),
          length.out = size
        ),
        levels = unique(sort(
          rep(
            x = c("group A", "group B", NA_character_),
            length.out = size
          )
        ))
      ),
      ord = factor(
        rep(
          x = c("level 1", "level 2", NA_character_),
          length.out = size
        ),
        levels = unique(sort(
          rep(
            x = c("level 1", "level 2", NA_character_),
            length.out = size
          )
        )),
        ordered = TRUE
      )
    )
  } else {
    switch(bin_type,
      log = rep(x = c(TRUE, FALSE), length.out = size),
      int = rep(x = c(0L, 1L), length.out = size),
      chr = rep(x = c("item:A", "item:B"), length.out = size),
      fct = factor(rep(x = c("group A", "group B"), length.out = size),
        levels = unique(sort(
          rep(
            x = c("group A", "group B"),
            length.out = size
          )
        ))
      ),
      ord = factor(
        rep(
          x = c("level 1", "level 2"),
          length.out = size
        ),
        levels = unique(sort(
          rep(
            x = c("level 1", "level 2"),
            length.out = size
          )
        )),
        ordered = TRUE
      )
    )
  }
}
facet_maker <- function(facet_type, size, lvls, missing = FALSE) {
  if (isTRUE(missing)) {
    switch(facet_type,
      chr = chr_maker(size = size, lvls = lvls, missing = TRUE),
      fct = fct_maker(size = size, lvls = lvls, missing = TRUE),
      ord = ord_maker(size = size, lvls = lvls, missing = TRUE)
    )
  } else {
    switch(facet_type,
      chr = chr_maker(size = size, lvls = lvls, missing = FALSE),
      fct = fct_maker(size = size, lvls = lvls, missing = FALSE),
      ord = ord_maker(size = size, lvls = lvls,  missing = FALSE)
    )
  }
}
col_maker <- function(col_type, size, ..., missing) {

  make_cols <- function(col_type, size, ..., missing) {
    switch(col_type,
     log = log_maker(size = size, missing = missing),
     int = int_maker(size = size, missing = missing),
     dbl = dbl_maker(size = size, missing = missing),
     chr = chr_maker(size = size, missing = missing, ...),
     fct = fct_maker(size = size, missing = missing, ...),
     ord = ord_maker(size = size, missing = missing, ...))
    }

  cols_list <- purrr::map(
    .x = col_type,
    .f = make_cols,
    size = size,
    missing = missing,
    ...)

  col_nms <- paste0(janitor::make_clean_names(col_type), "_var")

  cols_tbl_list <- purrr::map(
    .x = cols_list,
    .f = tibble::as_tibble)

  cols_tbl <- purrr::list_cbind(cols_tbl_list, size = size) |>
                  suppressMessages() |>
                  suppressWarnings()

  cols_tbl <- purrr::set_names(cols_tbl, col_nms)

  return(cols_tbl)

}
```

```{r pkgs, message=FALSE, warning=FALSE}
library(utap)
library(lobstr)
```

## Function tree

The `lobstr` package provides the abstract syntax tree function, which I like to use for creating function trees. For example, below is the tree for `select_column_class()`:

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    select_column_class(
        col_class() # nested (not defined in roxygen2)
      )
)
```

```{verbatim}
#| eval: false
#| echo: true
█─select_column_class 
└─█─col_class
```


This function can be demonstrated using data with some of the test helpers in `tests/testthat/helper.R`

```{r test_data}
#| eval: true
#| echo: true
test_data <- tibble::tibble(
    list_var = list(fct_vec = fct_maker(size = 3), 
                    ord_vec = ord_maker(size = 3),
                    chr_vec = chr_maker(size = 3),
                    log_vec = log_maker(size = 3),
                    int_vec = int_maker(size = 3),
                    dbl_vec = dbl_maker(size = 3)),
    log_var = log_maker(size = 6, missing = TRUE),
    int_var = int_maker(size = 6, missing = TRUE),
    dbl_var = dbl_maker(size = 6, missing = TRUE),
    facet_var = facet_maker(facet_type = "fct", 
                            lvls = 4, missing = TRUE, 
                            size = 6),
    bin_var = bin_maker(bin_type = "log", 
                        missing = TRUE, size = 6))
```

```{r}
#| eval: false
#| echo: true
test_data
```

```{r}
#| eval: true
#| echo: false
knitr::kable(test_data)
```


### select_column_class()

```{r}
#| eval: false
#| echo: true
select_column_class(
  df = test_data, 
  class = "log")

select_column_class(
  df = test_data, 
  class = c("int", "list"))
```



```{r}
#| eval: true
#| echo: false
select_column_class(
  df = test_data, 
  class = "log") |> knitr::kable()
```

```{r}
#| eval: true
#| echo: false
select_column_class(
  df = test_data, 
  class = c("int", "list")) |> knitr::kable()
```


### get_column_class()

`select_column_class()` is called from within `get_column_class()`, which is a wrapper function that includes the `return_tbl` argument for returning the columns as a named vector. 

```{r}
#| eval: false
#| echo: true
lobstr::ast(
  get_column_class(
    select_column_class(
      col_class() # nested (not defined in roxygen2)
      )
    )
)
```

```{verbatim}
#| eval: false
#| echo: true
█─get_column_class 
└─█─select_column_class 
  └─█─col_class 
```

```{r}
#| eval: false
#| echo: true
get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = TRUE)

get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = FALSE)
```

```{r}
#| eval: true
#| echo: false
get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = TRUE) |> 
  knitr::kable()
```

```{r}
#| eval: true
#| echo: false
get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = FALSE) 
```


### pull_[type]_cols()

All of the `pull_[type]_cols()` functions use the `get_column_class()` to return the initial set of column names. 

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    get_column_class(
      select_column_class(col_class()),
    pull_binary_cols(),
    pull_facet_cols(),
    pull_cat_cols(),
    pull_numeric_cols())
  )
```

```{verbatim}
#| eval: false
#| echo: true
█─get_column_class 
├─█─select_column_class 
│ └─█─col_class 
├─█─pull_binary_cols 
├─█─pull_facet_cols 
├─█─pull_cat_cols 
└─█─pull_numeric_cols
```


### pull_binary_cols()

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    pull_binary_cols(check_binary_vec(
                          check_log_binary(),
                          check_int_binary(),
                          check_fct_binary()),
                     make_binary_vec()))
```

```{verbatim}
#| eval: false
#| echo: true
█─pull_binary_cols 
├─█─check_binary_vec 
│ ├─█─check_log_binary 
│ ├─█─check_int_binary 
│ └─█─check_fct_binary 
└─█─make_binary_vec 
```

```{r}
pull_binary_cols(df = test_data)
```


### pull_facet_cols()

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    pull_facet_cols(check_facet_vec(
                          check_chr_facet(), 
                          check_fct_facet()),
                      make_facet_vec())
)
```

```{verbatim}
#| eval: false
#| echo: true
█─pull_facet_cols 
├─█─check_facet_vec 
│ ├─█─check_chr_facet 
│ └─█─check_fct_facet 
└─█─make_facet_vec 
```

```{r}
pull_facet_cols(df = test_data)
```

