---
title: "utap"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{utap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "\t",
  eval = TRUE
)
source("../tests/testthat/helper.R")
```

```{r pkgs, message=FALSE, warning=FALSE}
library(utap)
library(lobstr)
```

## Function tree

The `lobstr` package provides the abstract syntax tree function, which I like to use for creating function trees. For example, below is the tree for `select_column_class()`:

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    select_column_class(
        col_class() # nested (not defined in roxygen2)
      )
)
```

```{verbatim}
#| eval: false
#| echo: true
█─select_column_class 
└─█─col_class
```


This function can be demonstrated using data with some of the test helpers in `tests/testthat/helper.R`

```{r test_data}
#| eval: true
#| echo: true
test_data <- tibble::tibble(
    list_var = list(fct_vec = fct_maker(size = 3), 
                    ord_vec = ord_maker(size = 3),
                    chr_vec = chr_maker(size = 3),
                    log_vec = log_maker(size = 3),
                    int_vec = int_maker(size = 3),
                    dbl_vec = dbl_maker(size = 3)),
    log_var = log_maker(size = 6, missing = TRUE),
    int_var = int_maker(size = 6, missing = TRUE),
    dbl_var = dbl_maker(size = 6, missing = TRUE),
    facet_var = facet_maker(facet_type = "fct", 
                            lvls = 4, missing = TRUE, 
                            size = 6),
    bin_var = bin_maker(bin_type = "log", 
                        missing = TRUE, size = 6))
```

```{r}
#| eval: false
#| echo: true
test_data
```

```{r}
#| eval: true
#| echo: false
knitr::kable(test_data)
```


### select_column_class()

```{r}
#| eval: false
#| echo: true
select_column_class(
  df = test_data, 
  class = "log")

select_column_class(
  df = test_data, 
  class = c("int", "list"))
```



```{r}
#| eval: true
#| echo: false
select_column_class(
  df = test_data, 
  class = "log") |> knitr::kable()
```

```{r}
#| eval: true
#| echo: false
select_column_class(
  df = test_data, 
  class = c("int", "list")) |> knitr::kable()
```


### get_column_class()

`select_column_class()` is called from within `get_column_class()`, which is a wrapper function that includes the `return_tbl` argument for returning the columns as a named vector. 

```{r}
#| eval: false
#| echo: true
lobstr::ast(
  get_column_class(
    select_column_class(
      col_class() # nested (not defined in roxygen2)
      )
    )
)
```

```{verbatim}
#| eval: false
#| echo: true
█─get_column_class 
└─█─select_column_class 
  └─█─col_class 
```

```{r}
#| eval: false
#| echo: true
get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = TRUE)

get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = FALSE)
```

```{r}
#| eval: true
#| echo: false
get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = TRUE) |> 
  knitr::kable()
```

```{r}
#| eval: true
#| echo: false
get_column_class(
  df = test_data, 
  class = c("int", "dbl"),
  return_tbl = FALSE) 
```


### pull_[type]_cols()

All of the `pull_[type]_cols()` functions use the `get_column_class()` to return the initial set of column names. 

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    get_column_class(
      select_column_class(col_class()),
    pull_binary_cols(),
    pull_facet_cols(),
    pull_cat_cols(),
    pull_numeric_cols())
  )
```

```{verbatim}
#| eval: false
#| echo: true
█─get_column_class 
├─█─select_column_class 
│ └─█─col_class 
├─█─pull_binary_cols 
├─█─pull_facet_cols 
├─█─pull_cat_cols 
└─█─pull_numeric_cols
```


### pull_binary_cols()

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    pull_binary_cols(check_binary_vec(
                          check_log_binary(),
                          check_int_binary(),
                          check_fct_binary()),
                     make_binary_vec()))
```

```{verbatim}
#| eval: false
#| echo: true
█─pull_binary_cols 
├─█─check_binary_vec 
│ ├─█─check_log_binary 
│ ├─█─check_int_binary 
│ └─█─check_fct_binary 
└─█─make_binary_vec 
```

```{r}
pull_binary_cols(df = test_data)
```


### pull_facet_cols()

```{r}
#| eval: false
#| echo: true
lobstr::ast(
    pull_facet_cols(check_facet_vec(
                          check_chr_facet(), 
                          check_fct_facet()),
                      make_facet_vec())
)
```

```{verbatim}
#| eval: false
#| echo: true
█─pull_facet_cols 
├─█─check_facet_vec 
│ ├─█─check_chr_facet 
│ └─█─check_fct_facet 
└─█─make_facet_vec 
```

```{r}
pull_facet_cols(df = test_data)
```

